comgen-outline

### 2017-08-29 paper/comgen-paper.Rmd
add a note
rewrite introduction to incorporate generics literature and tighten

Knowing that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to trhiving in our open-ended world.
Yet such knowledge can be difficult to acquire: Relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding the properties of lightning).
It thus is crucial that language allow us to communicate such *generalizations* to each other.
By sharing generalizations we thrive collectively without individually needing to taste each potentialy-poisonous plant or witness enough lightning strikes.
Furthermore, sharing generalizations from one generation to the next may be a crucial mechanism behind the cultural ratchet \ndg{cite. or remove if you don't like it here}.

Generalizations can be acquired from the world by drawing inductive inferences from observations [@Baldwin1993; @Xu2008; @Dewar2010]. \ndg{cite older stuff. eg hume}
Cooperative teachers can facilitate inductive learning by supplying helpful observations to the learner [@Csibra2009; @Tomasello1999; @Butler2012].
At around a child's first birthday, however, the learner gains access to a new, human-unique source of data: language.
Minimally, language fosters generalizations by bringing attention to objects being of the same kind through the use of labeling [e.g., "This is a dog"; @Gelman1986; @Gopnik2000; @Graham2004; @Markman1989; @Schulz2008].
Objects have many properties though; when speakers attach properties to kind-labels (e.g., "Dogs bark"), listeners hear a generalization.
\ndg{this par doesn't really do much}

Generalizations in language are often studied in *generic language* or generics.
Generic language is ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
Around the age of two and a half, children begin to understand that generic language refers to categories and conveys information that extends beyond the present context [@Cimpian2008].
It is believed that every language can express generic meaning [@Behrens2005; @Carlson1995], and such generalizations in language are thought to be central to the growth of conceptual knowledge [@Gelman2004].
<!-- *Causal language* (e.g., "The block *makes* it [the machine] go.") faciliates more abstract forms of causal reasoning [@Bonawitz2010]. -->
Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010].

Generic statements ascribing properties to categories are just one kind of generalization found in language. 
Individual entities can also be talked about in generalization, using so-called *habitual language* (e.g., "Mary swims after work.").
<!-- Recent evidence suggests talking about the other in generalization (using so-called *generic You*) can help make meaning from difficult situations   [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->
Communicating about causality can also take the form of communicating a generalization about a causal event (*causal language*; e.g., "Drinking moonshine makes you go blind.").
What these statements have in common is that they convey generalizations, or exhibit *genericity* [@Carlson1977; @Carlson1995].

@@

\ndg{prev three pars are floppy. tighten.
There haven't been many formal models of generic language, none have made quantitative predictions. Say why this is important.
To build one, we need to firsts say what a generalization is. We use the Bayesian idea that an inductive generalization can be understood in terms if probbilities.
Then we need to say how these probabilities are communicated. Use RSA...}

@@

The major issue in formalizing generalizations in language is determining what makes such statements true or false.
When we hear "Dogs bark", it feels like a universal statement, like *All* dogs bark.
But the property doesn't apply to all dogs: Basenji dogs are barkless^[
http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html
].
So, the generalization should signal something weaker like *Most* dogs bark.
<!-- Indeed, many generalizations seem to have the quantificational force of *most*. -->
However, *Mosquitos carry malaria* is intuitively a true statement, despite less than 1\% of moquitos actually carrying malaria.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only quantifier that would permit such a statement would be "Some" (e.g., *some mosquitos carry malaria*).
But weakening the generalization to *some* goes too far: Half of robins are female, but saying *Robins are female* seems odd.
That same half of robins, however, also lay eggs, and *Robins lay eggs* is again intuively reasonable.

@@

Rational models of communication [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016] have had a lot of recent success in formalizing complex linguistic phenomena ranging from politeness [@Yoon2016] and hyperbole [@Kao2014] to metaphor [@Kao2014] and vagueness [@Lassiter2013; @Qing2014]. 
Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their communicative goals. 
Heretofore, however, the fragments of language under investigation have been limited to utterances that convey information about a specific entity or situation, about the here and now.
Meanwhile, the language of generalizations don't make reference to entities in the world or concrete situations. 
They refers to concepts in the head and have proved tricky to formalize.

In this paper, we explore the hypothesis that the core meaning of a generic statement and other language for conveying generalizations is *simple but underspecified*.
We formalize this hypothesis in a probabilistic model of language understanding, and derive predictions for the endorsements of such statements by considering whether or not a speaker would say the generalization to a listener.
We show that this decision depends in systematic ways on the prior knowledge of the listener.

The paper is organized as follows. 
In Section 1, we describe the Bayesian framework for generalization and define the scales along which generalization is measured in our three case studies. 
In Section 2, we formalize the hypothesis that the core of meaning of a generalization is *simple but underspecified* and incorporate this *uncertain semantics* into a framework for pragmatic language understanding -- the Rational Speech Act framework.

To test our model, we examine endorsements of generalizations in three domains: categories, events, and causes.
Our model makes predictions about how endorsements should vary according to (i) the listener's world knowledge prior and (ii) the speaker's intended message (or, belief distribution) that she seeks to communicate. 
In Section 3, we describe experiments the measure both quantities and relate them to the endorsements of generalizations about categories, or *generic language*. 
In Section 4, we describe experiments that manipulate the speaker's intended message in communicating generalizations about events, or *habitual language*.
In Section 5, we describe experiments that manipulate both the listener's relevant prior knowledge and the speaker's intended message in communicating generalizations about causes, or *causal language*. 
Across all of our experimental settings, we find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements (cumulative $r^2(N) = 0.XX$), where simpler models fall short.
In the final section of the paper, we discuss how our model begins the process of unifying the extant psychological, linguistic, and philosophical literatures on *genericity*.
Our theory also opens up a number of new directions for understanding generalizations in language and building models that learn from language like people.

@@

# Generalization from observations

@@

We will discuss communicating generalizations in the context of generalizations about categories (i.e., *generic language*), though the arguments will extend to events and causes as well, following the definitions of $h$ described in the previous section. 
Generics express a relation between a kind K (e.g., \textsc{dogs}) and a property F (e.g., \textsc{is friendly}), such that the property can also be said to be applicable to an individual (i.e., "My dog is friendly"). 
Bare plural statements (e.g., *Dogs are friendly*) strongly tend to yield a generic meaning, though other forms can express such a meaning [e.g., *A mongoose eats snakes.*; @Carlson1977] .

Given that generics express a property that can be applied to individuals, it seems intuitive that what makes the statement true or false would correspond to the number of individuals who hold the property.
At first glance, generics convey that *most* have the property, such as *Most dogs are friendly*.
*Robins lay eggs* is true, though, despite the fact that only female robins lay eggs, which is debatable whether or not that counts as *most robins*. 
If generic statements can be true by restricting themselves to only be describing females, why does *Robins are female* seem like a weird statement to make?
Finally, *Mosquitos carry malaria* is intuitively true despite malaria only being present in a tiny fraction of a percentage of mosquitos.
It seems that any hard condition on the prevalence of the feature in the category (e.g., how many robins lay eggs) would violate intuitions.

These observations have led to proposals that prevalence or probability is only peripherally related to generic meaning. 
**Conceptual accounts** of generics emphasize the structure of generic knowledge [@Prasada2000], and view generic utterances as a special way of expressing mental relationships between kinds and properties [@Leslie2008; @Prasada2012]. 
From this perspective, the statement *Bishops move diagonally* is true not because most bishops move diagonally, but rather those are the rules of chess.
The most influential account in psychology is credited to @Leslie2007, who proposes that generics express a cognitively primitive, *default generalization*.

The *generics as default* view argues that general cognitive principles explain why certain generics are true or not.
*Mosquitos carry malaria* conveys a dangerous or striking property, which is useful to know about if you are trying to survive in a hostile world.
This theory thus predicted that generics that convey striking properties would be true even if they were relatively rare (e.g., *Sharks attack swimmers*), which was empirically corroborated [@Cimpian2010; @Prasada2013].
Other generics convey properties that are *characteristic* of a kind, such as *Robins lay eggs*, argued to have a *principled connection* between kind and property [@Prasada2006; @Prasada2013].

Yet, generics are not limited to conveying rich relations. 
Arbitrary relationships can also be expressed (e.g., *Barns are red*; *Ravens are bigger than toasters*) which seem only to be true because those are the statistics of the world [@Nickel2008; @Nickel2016; @Sterken2015].
The *generics as default* view must then be supplemented with the principle that, in certain situations,  statistics can influence generic endorsement.
It is thus argued that there are qualitatively different kinds of generic statements (i.e., "striking", "characteristic", "statistical"), each with their own conditions on what makes them true.
This diversity is interested, but what is common across generics?
Is there some core meaning to *genericity*, a general unified principle that explains what makes generic statements true or false?

We propose that generic language and other forms of genericity information convey information about probabilities that underly generalization and prediction.
We posit that the conceptual factors of the kind discussed by Prasada, Leslie, and colleagues influence the meaning of a generic statement via their influence on probabilities. 
Bayesian cognitive models have been shown to represent flexible, structured knowledge of the world and formalize the inferences that people draw using such conceptual representations [e.g., @Kemp2008; @Goodmanconcepts].
Thee basic currency of inference in these models is probability.
In generalizations in language convey information about probabilties, then a relatively simple semantic theory (e.g., a threshold on probability) may be enough to formalize the core meaning of a generic statement $u_{gen}$:
<!-- In the case of generic language, we posit that conceptual factors may influence the meaning of a generic statement via its influence on subjective probability. -->
<!-- If that's true, a simple, semantic theory  -->

>>>>

Knowing that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be necessary to thrive in our open-ended world.
Yet such knowledge can be difficult to acquire: Relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
It thus is crucial that language allow us to communicate such *generalizations* to each other.
By sharing generalizations we flourish collectively without individually needing to taste each potentially-poisonous plant or witness enough lightning strikes.
Furthermore, communicating generalizations from one generation to the next may be a foundational mechanism behind the faithful transmission of knowledge necessary for cumulative cultural evolution [@Tomasello1999; @Henrich2015].

<!-- Generalizations can be acquired from the world by drawing inductive inferences from observations [@Baldwin1993; @Xu2008; @Dewar2010]. \ndg{cite older stuff. eg hume} -->
<!-- Cooperative teachers can facilitate inductive learning by supplying helpful observations to the learner [@Csibra2009; @Tomasello1999; @Butler2012]. -->
<!-- At around a child's first birthday, however, the learner gains access to a new, human-unique source of data: language. -->
<!-- Minimally, language fosters generalizations by bringing attention to objects being of the same kind through the use of labeling [e.g., "This is a dog"; @Gelman1986; @Gopnik2000; @Graham2004; @Markman1989; @Schulz2008]. -->
<!-- Objects have many properties though; when speakers attach properties to kind-labels (e.g., "Dogs bark"), listeners hear a generalization. -->
<!-- \ndg{this par doesn't really do much} -->

Generalizations in language are often studied in *generic language* (or, generics): simple statements that convey generalizations about categories [e.g., "Dogs are friendly"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008].
In contrast to *particular statements* which describe individual entities or situations (e.g., "Rufus is friendly") and can be verified through observation, generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dogs}) and convey information that extends beyond the present context, a fact which two-and-a-half-year-olds recognize [@Cimpian2008].
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
Generalizations are not limited to generics, however; they can be found in statements about individuals (*habitual language*; e.g., "Mary swims after work.") as well as causal events (*causal language*; e.g., "Staring at the sun makes you go blind.").
It is believed that every language can express generalizations [@Behrens2005; @Carlson1995] and that such language is central to the growth of conceptual knowledge [@Gelman2004].

<!-- Generic statements ascribing properties to categories are just one kind of generalization found in language.  -->
<!-- Individual entities can also be talked about in generalization, using so-called  -->
<!-- <!-- Recent evidence suggests talking about the other in generalization (using so-called *generic You*) can help make meaning from difficult situations   [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->
<!-- Communicating about causality can also take the form of communicating a generalization about a causal event  -->
<!-- What these statements have in common is that they convey generalizations, or exhibit *genericity* [@Carlson1977; @Carlson1995]. -->

Despite their ubiquity and relative simplicity, there have been few formal models of generic language and none that have made precise, quantitative predictions about observable behavior.
A formal model in this domain is advantageous to make explicit the relevant assumptions of a psychological or linguistic hypothesis for predicting human behavior; this, in turn, makes the theory more falsifiable. 
The major challenge in formalizing generalizations in language is determining what makes such statements true or false.
When we hear "Dogs bark", it feels like a universal statement, like *All* dogs bark.
But the property doesn't apply to all dogs: Basenji dogs are barkless^[
http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html
].
The generalization could signal something weaker like *Most* dogs bark.
<!-- Indeed, many generalizations seem to have the quantificational force of *most*. -->
However, less than 1\% of moquitos actually carry malaria and still "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only quantifier that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to *some* goes too far: Half of robins are female, but saying "Robins are female" seems odd.
That same half of robins (i.e., the females), however, also lay eggs, and "Robins lay eggs" is intuively reasonable.

These observations have led to proposals that the statistics of the feature (e.g., the prevalence or probability of a member of category having a property) is only incidentally related to generic meaning. 
**Conceptual accounts** of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012]. 
From this perspective, the statement "Bishops move diagonally" is true not because most bishops move diagonally, but rather because those are the rules of chess.

With this perspective, @Leslie2007 proposes generics express a cognitively primitive, default generalization. 
The *generics as default* view argues that certain principles explain how generics become true or false.
"Mosquitos carry malaria" conveys a dangerous or *striking property*, which is useful to know about if you are trying to survive in a hostile world.
This theory correctly predicted that generics that convey striking properties would be true even if the property were relatively rare [e.g., "Sharks attack swimmers"; @Cimpian2010; @Prasada2013].
"Robins lay eggs" can be thought of as conveying a *characteristic property*, which permits its felicity despite only 50\% having the feature [@Prasada2006; @Prasada2013].

Yet, generics are not limited to conveying rich, conceptual relations. 
Arbitrary relationships can also be expressed with generics (e.g., "Barns are red"; "Ravens are bigger than toasters") which seem only to be true because those are the statistics of the world [@Nickel2008; @Nickel2016; @Sterken2015].
The *generics as default* view must then be supplemented with the principle that, in certain situations, statistics influence generic endorsement.
The diversity of generic statements highlighted by this view is interesting, but insofar as generics are a single, unified phenomenon, there should be a general, unified principle that explains what makes the statements true or false.
<!-- It is thus argued that there are qualitatively different kinds of generic statements (i.e., "striking", "characteristic", "statistical"), each with their own conditions on what makes them true. -->
<!-- This diversity is interesting, but what is common across generics? -->
<!-- Is there some core meaning to *genericity*, a general unified principle that explains what makes generic statements true or false? -->

We propose a formal, unified account of generalizations in language that can make precise, quantitative predictions about these simple and ubiquituous utterances.
We do so by first describing the much more well-studied form of generalization: generalization from observations. 
The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is a good model of human generalizations [@Shepard1987; @Tenenbaum2001].
We then formalize how these probabilities are communicated, using a rational model of communication --- the Rational Speech Act theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016].
Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals. 
Heretofore, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013].
The language of generalizations, however, does not make reference to entities in the world or concrete situations, but instead refers to concepts in the head.

Specifically, we explore the hypothesis that the core meaning of a generic statement and other language for conveying generalizations concerns subjective probability in a *simple but underspecified* way.
We formalize this hypothesis in a probabilistic model of language understanding, and derive predictions for the endorsements of such statements by considering whether or not a speaker would say the generalization to a listener.
We show that this decision depends in systematic ways on the prior knowledge of the listener.

The paper is organized as follows. 
In Section 1, we describe the Bayesian framework for generalization from observations, serving as the foundation for our semantic theory.
In Section 2, we formalize the hypothesis that the core of meaning of a generalization is *simple but underspecified* and incorporate this *uncertain semantics* into a framework for pragmatic language understanding -- the Rational Speech Act framework.

To test our model, we examine endorsements of generalizations in three domains: categories, events, and causes.
Our model makes predictions about how endorsements should vary according to (i) the listener's world knowledge prior and (ii) the speaker's state of knowledge (or, belief distribution) that she seeks to communicate. 
In Section 3, we describe experiments the measure both quantities and relate them via our model to the endorsements of generalizations about categories, or *generic language*. 
In Section 4, we describe experiments that manipulate the speaker's state of knowledge in communicating generalizations about events, or *habitual language*.
In Section 5, we describe experiments that manipulate both the listener's relevant prior knowledge and the speaker's state of knowledge in communicating generalizations about causes, or *causal language*. 
Across all of our experimental settings, we find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements (cumulative $r^2(N) = 0.XX$), where simpler models fall short.
In the final section of the paper, we discuss how our model begins the process of unifying the extant psychological, linguistic, and philosophical literatures on *genericity*.
Our theory also opens up a number of new directions for understanding generalizations in language and building models that learn from language in the way that people learn from language.

<!-- (2) It also allows us to consider sources of gradience in behavior, determining whether or not we should allocate it to *noise* in the observed data or to *uncertainty* in the head of the language user. -->
<!-- *Causal language* (e.g., "The block *makes* it [the machine] go.") faciliates more abstract forms of causal reasoning [@Bonawitz2010]. -->

<!-- \ndg{There haven't been many formal models of generic language, none have made quantitative predictions. Say why this is important. -->
<!-- To build one, we need to firsts say what a generalization is. We use the Bayesian idea that an inductive generalization can be understood in terms if probbilities. -->
<!-- Then we need to say how these probabilities are communicated. Use RSA...} -->

<!-- Rational models of communication have had a lot of recent success in formalizing complex linguistic phenomena ranging from politeness [@Yoon2016] and hyperbole [@Kao2014] to metaphor [@Kao2014] and vagueness [@Lassiter2013; @Qing2014].  -->

@@

@@

@@

@@

# Generalization from Observations

@@

Generics express a relation between a kind K (e.g., \textsc{dogs}) and a property F (e.g., \textsc{is friendly}), such that the property can also be said to be applicable to an individual (i.e., "My dog is friendly"). 
Bare plural statements (e.g., "Dogs are friendly") tend to yield a generic meaning, though other forms can express such a meaning [e.g., "A mongoose eats snakes."; @Carlson1977] .
Given that generics express a property that can be applied to individuals, it seems intuitive that what makes the statement true or false would correspond to the number of individuals who hold the property.
At first glance, generics convey that *most* have the property, such as "Most dogs are friendly".
"Robins lay eggs" is true, though, despite the fact that only female robins lay eggs.
If generic statements can be true by restricting themselves to only be describing females, why does *Robins are female* seem like a weird statement to make?
Finally, "Mosquitos carry malaria" is intuitively true despite malaria only being present in a tiny fraction of a percentage of mosquitos.
It seems that any hard condition on the prevalence of the feature in the category (e.g., how many robins lay eggs) would violate intuitions.

We propose that generic language and other forms of genericity information convey information about probabilities that underly generalization and prediction.
We posit that the conceptual factors of the kind discussed by Prasada, Leslie, and colleagues influence the meaning of a generic statement via their influence on probabilities. 
Bayesian cognitive models have been shown to represent flexible, structured knowledge of the world and formalize the inferences that people draw using such conceptual representations [e.g., @Kemp2008; @Goodmanconcepts].
The basic currency of inductive reasoning in these models is probability.
If generalizations in language convey information about probabilties, then a relatively simple semantic theory (e.g., a threshold on probability) may be enough to formalize the core meaning of a generic statement $u_{gen}$:

\ndg{move all this background about generics earlier, so that this section can go straight to our model.}

We will discuss communicating generalizations in the context of generalizations about categories (i.e., *generic language*), though the arguments will extend to events and causes as well, following the definitions of $h$ described in the previous section.

<!-- In the case of generic language, we posit that conceptual factors may influence the meaning of a generic statement via its influence on subjective probability. -->
<!-- If that's true, a simple, semantic theory  -->
